---
title: "SpaDES Module GitHub Structure Guide"
author: "Project Reference Document"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
description: >
  This document defines the standard conventions for storing SpaDES modules on
  GitHub. It is intended to be read by Claude Code as a reference when auditing,
  restructuring, or scaffolding an existing SpaDES project to conform to
  community standards.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

---

# Purpose

This document serves as a **machine-readable conventions reference** for Claude Code.
When modifying an existing SpaDES project, use the rules and structures defined
here to:

- Audit the project layout and flag deviations
- Scaffold missing files or directories
- Rename repositories or modules that violate naming conventions
- Generate or update `global.R`, module scripts, and helper files
- Suggest or apply `git submodule` wiring for tandem module suites

---

# 1. Core Naming Convention

## Rule
> The GitHub repository name **must exactly match** the module name declared
> inside the module's `defineModule()` call.

## Why It Matters
`SpaDES::downloadModule()` and `SpaDES.install::installModules()` resolve modules
by matching the repo name to the module name. A mismatch will cause download
failures.

## Check to Perform
```{r naming-check}
# Claude Code: In the target project, verify that for each module directory:
# 1. The folder name matches the .R script name inside it
# 2. The name() field inside defineModule() matches both of the above

# Example of a correctly named module script header:
defineModule(sim, list(
  name = "MyModule",           # <-- must match repo name and folder name
  description = "Does X",
  keywords = c("SpaDES"),
  authors = person("First", "Last", email = "email@example.com",
                   role = c("aut", "cre")),
  childModules = character(0),
  version = list(MyModule = "1.0.0"),
  timeframe = NA,
  timeunit = "year",
  citation = list(),
  documentation = list(),
  reqdPkgs = list(),
  parameters = bindrows(),
  inputObjects = bindrows(),
  outputObjects = bindrows()
))
```

---

# 2. GitHub Organization Structure

## Convention
All related modules in a tandem suite are stored under a single **GitHub
Organization**, not scattered across personal accounts.

## Expected Layout on GitHub
```
https://github.com/{OrgName}/
├── ModuleA/        # one repo per module
├── ModuleB/
├── ModuleC/
└── myProject/     # separate controller/orchestration repo
```

## Actions for Claude Code
- If modules are found under a personal account (`github.com/username/`) rather
  than an org, flag this and note the recommended migration path.
- Do **not** automatically rename or move repos; flag for human review.

---

# 3. Individual Module Repository Layout

## Required Structure
Each module repository must contain the following layout. Claude Code should
create any missing files/directories listed here.

```
ModuleA/                     # repo root — same name as the module
├── ModuleA.R                # primary module script (required)
├── ModuleA.Rmd              # documentation/vignette (required)
├── R/                       # helper functions sourced by the module (create if absent)
│   └── helpers.R            # example helper (optional content)
├── data/                    # input data directory (create if absent)
│   └── .gitkeep             # placeholder so git tracks the empty dir
├── tests/                   # test directory (create if absent)
│   └── testthat/
│       └── test-moduleA.R   # unit tests
└── NEWS.md                  # changelog (create if absent)
```

## Files That Must NOT Be Committed
Ensure `.gitignore` contains at minimum:

```
# SpaDES outputs and caches
outputs/
cache/
.RData
.Rhistory

# Large data files (use Git LFS or external storage instead)
data/*.csv
data/*.tif
data/*.rds
```

---

# 4. Project Controller Repository

## Purpose
When modules are designed **in tandem**, a separate **project controller repo**
orchestrates them. This repo does **not** contain module source code — it only
wires modules together and runs the simulation.

## Required Layout
```
myProject/
├── global.R            # top-level simInit() + spades() call (required)
├── modules/            # modules downloaded/checked out here
│   ├── ModuleA/
│   ├── ModuleB/
│   └── ModuleC/
├── inputs/             # simulation inputs
├── outputs/            # simulation outputs (gitignored)
├── .Rprofile           # project-level R options (optional)
└── README.md
```

## Minimum `global.R` Template
Claude Code may generate this file if it is absent:

```{r global-template}
# global.R — SpaDES project controller

library(SpaDES.core)
library(SpaDES.install)

# ── Paths ──────────────────────────────────────────────────────────────────
projectPath <- getwd()
modulePath  <- file.path(projectPath, "modules")
inputPath   <- file.path(projectPath, "inputs")
outputPath  <- file.path(projectPath, "outputs")
cachePath   <- file.path(projectPath, "cache")

# ── Download / install modules if not present ──────────────────────────────
installModules(
  modules = c("ModuleA", "ModuleB", "ModuleC"),
  modulePath = modulePath,
  account   = "YourGitHubOrg"   # <-- replace with actual org name
)

# ── Simulation parameters ──────────────────────────────────────────────────
times  <- list(start = 0, end = 10)
params <- list(.globals = list(verbose = TRUE))

modules <- list("ModuleA", "ModuleB", "ModuleC")

# ── Initialise and run ─────────────────────────────────────────────────────
mySim <- simInit(
  times      = times,
  params     = params,
  modules    = modules,
  paths      = list(
    modulePath = modulePath,
    inputPath  = inputPath,
    outputPath = outputPath,
    cachePath  = cachePath
  )
)

outputs(mySim) <- spades(mySim, debug = TRUE)
```

---

# 5. Tandem Module Versioning

## Convention
Each module repo uses **semantic version tags** so the project controller can pin
compatible versions.

### Tagging a release (run in each module repo)
```bash
git tag -a v1.0.0 -m "Release v1.0.0"
git push origin v1.0.0
```

## Git Submodules (Optional — for strict reproducibility)
If the project uses `git submodule` to pin module versions inside the controller
repo:

```bash
# Add each module as a submodule (run from project root)
git submodule add https://github.com/YourGitHubOrg/ModuleA modules/ModuleA
git submodule add https://github.com/YourGitHubOrg/ModuleB modules/ModuleB
git submodule add https://github.com/YourGitHubOrg/ModuleC modules/ModuleC

# Pin to a specific tag
cd modules/ModuleA && git checkout v1.0.0 && cd ../..
git add .gitmodules modules/
git commit -m "Pin ModuleA to v1.0.0"
```

### Claude Code action
- If `modules/` contains plain directories with no `.gitmodules` file, flag the
  lack of version pinning.
- If `.gitmodules` exists, verify each submodule URL matches the expected org.

---

# 6. Module Download Convention

SpaDES resolves modules using this URL pattern:

```
https://github.com/{account}/{moduleName}/archive/{version}.zip
```

## Validation Check for Claude Code
For each module in the project, confirm:

```{r download-check}
# Pseudocode — adapt to actual project structure
expected_modules <- c("ModuleA", "ModuleB", "ModuleC")
org              <- "YourGitHubOrg"

for (mod in expected_modules) {
  url <- paste0("https://github.com/", org, "/", mod)
  message("Expected repo URL: ", url)
  # Verify the local folder name matches `mod`
  local_path <- file.path("modules", mod)
  if (!dir.exists(local_path)) {
    warning("Module directory missing: ", local_path)
  }
}
```

---

# 7. Summary Checklist for Claude Code

When applied to an existing SpaDES project, Claude Code should verify and
remediate each item below. Items marked **(auto-fix)** can be created or
corrected automatically; items marked **(flag)** should be reported to the user
for manual action.

| # | Check | Action |
|---|-------|--------|
| 1 | Repo name matches `defineModule(name = ...)` | **(flag)** if mismatch |
| 2 | Each module has its own repository | **(flag)** if co-located |
| 3 | All tandem modules share a GitHub Org | **(flag)** if on personal accounts |
| 4 | `ModuleA/R/` directory exists | **(auto-fix)** create if absent |
| 5 | `ModuleA/data/` with `.gitkeep` exists | **(auto-fix)** create if absent |
| 6 | `ModuleA/tests/testthat/` exists | **(auto-fix)** create if absent |
| 7 | `.gitignore` covers outputs, cache, large data | **(auto-fix)** append missing entries |
| 8 | Controller repo has `global.R` | **(auto-fix)** generate from template in §4 |
| 9 | `modules/` in controller is populated or submoduled | **(flag)** if empty |
| 10 | Semantic version tags exist on module repos | **(flag)** if no tags found |

---

# 8. References

- [SpaDES.core documentation](https://spades-core.predictiveecology.org/)
- [SpaDES.install documentation](https://spades-install.predictiveecology.org/)
- [PredictiveEcology GitHub Organization](https://github.com/PredictiveEcology)
- [SpaDES book](https://spades4dummies.predictiveecology.org/)
